Do syntactic and semantic similarity lead to interference effects?\\ Evidence from self-paced reading and event-related potentials using German

### 2024-11-30 paper/revision2/ResponseToReviewers.tex

\subsubsection*{Response to R2.1}
We agree with the reviewer that the presentation and discussion of the SPR results was misleading and have revised it accordingly. We have revised Figure 3 to remove the focus on the critical region (see below). At the beginning of the Results section, we now anticipate our interpretation of the results and state that we believe that the distractors led to encoding interference. We also state that the results at the critical (and spill-over) region are potentially confounded because it is not possible to disentangle the long-lasting encoding effect from retrieval effects potentially arising in the critical region.

\setcounter{figure}{2}

\begin{sidewaysfigure}[h]
    \caption{Self-paced reading times with 95\% confidence intervals. Panel A and B show the pooled reading times across the whole sentence; separately for high (A) and low (B) syntactic interference due to differing sentence structure. \textcolor{blue}{Panel C shows the reading times of the sentence focusing on the regions between the distractor and critical verb for all conditions.}}
    \label{fig:whole_sentence}
    \centering
    \includegraphics[width=\textwidth]{Pandora_all_wholesentence_pooled_zoom.jpg}
\end{sidewaysfigure}
\clearpage

The reviewer suggests that encoding interference confounded the reading times of all regions after the distractor and argues that this makes the results uninterpretable. The implication seems to be that the patterns observed at the pre-critical, critical, and spillover regions should not be reported.
 However, as we had already reported in the discussion of the SPR results, the two published studies that used the present design \citep{vandyke07, mertzen} show clear indications that these regions of interest were affected as well, possibly by encoding interference. Thus, across the three studies (Van Dyke et al., Mertzen et al., and ours) one consistent pattern is that the observed effects could be driven at least partly by encoding interference. We feel that this consistent pattern across studies in the three regions of interest is worth reporting.

@@

We discuss this on page \pageref{previous_encoding}:

\begin{quote}
    \Paste{previous_encoding}
\end{quote}

To summarize, one valuable finding in our study is the replication of the pattern found in the previous studies, which is attributable to encoding interference. Ours is the only study among these three to prominently raise the possibility that the reading times of the critical region may be confounded by encoding interference--we believe the reviewer would agree with this discussion. To ensure comparability to the previous studies, we consider it very important to report the reading times effects in the same regions which the previous studies reported, namely: the pre-critical, critical and spill-over region. This seems necessary to us despite the possible confound discussed above that the reviewer justifiably raises.

We have revised the Results section to be more concise, starting on page \pageref{revised_SPR}:

### 2024-12-01 paper/revision2/elsarticle-template-harv.tex
final edits

\subsection*{Discussion}
\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{However, reading times started to differ before this critical dependency could be formed. Starting at the distractor which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors). Because the distractor differed between conditions, we did not analyze distractor reading times. Instead, to investigate the difference starting after the distractor was read, we investigate the pre-critical region, which was identical across conditions and preceded the critical verb. The pre-critical region serves as a proxy for any encoding interference effects that arose after the distractor was read. Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region. In contrast, there was evidence against syntactic interference and an interaction in that region. We refrain from interpreting effects in the critical and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes, i.e., memory retrieval. The reading times differences at the distractor and the pre-critical region are discussed below.}}

@@

\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference. T}he most likely explanation is encoding interference (\cite{Oberauer_Kliegl_2006}, but see the potential confounds discussed above). The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.

>>>>

\subsection*{Discussion}
\Copy{SPR_discussion_revised}{We presented the to-date largest-sample self-paced reading study that \textcolor{blue}{aimed to investigate} the use of syntactic and semantic features during subject-verb dependency formation. \textcolor{blue}{However, reading times started to differ before this critical dependency could be formed. Starting at the distractor, which intervened between subject and verb, high semantic interference conditions (animate distractors) were read slower than low semantic interference conditions (inanimate distractors). Because the distractor differed between conditions, we did not analyze distractor reading times. Instead, to investigate the difference starting after the distractor was read, we investigate the pre-critical region, which was identical across conditions and preceded the critical verb. The pre-critical region serves as a proxy for any encoding interference effects that arose after the distractor was read. Bayes factors provided extremely strong evidence for a semantic interference effect in the pre-critical region. In contrast, there was evidence against syntactic interference and an interaction in that region. We refrain from interpreting effects in the critical and spill-over region, because it is unclear whether they were caused by the difference starting at the distractor or other processes, i.e., memory retrieval. The reading times differences at the distractor and the pre-critical region are discussed below.}}

@@

\textcolor{blue}{Since the reading times started to differ long before the critical verb which is assumed to be the locus of retrieval in the cue-based retrieval framework,  retrieval interference cannot be the cause of this difference. T}he most likely explanation is encoding interference \cite[e.g.,][but see the potential confounds discussed above]{Oberauer_Kliegl_2006,smith2021encoding}. The increased effort of encoding and subsequently maintaining representations of three different animate noun phrases (the introduction noun, the subject and the animate distractor) is the most likely reason for the slowdown in high vs.\ low semantic interference conditions \citep[for similar findings, see e.g., ][]{lago_etal_2021, ness2019, ness2017, kush_etal_2015, gordon02}.
